Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2020-04-06 paper/notes.tex
f is a function of rho, not directly of x.
Further tweaking of the meaning of the vvarious symbols and of what f() actually is and represents

On arriving at the tectum, axon growth cones must determine their direction of
movement. The model encapsulates the quantitative hypotheses that growth cone
movement is a combination of diffusion and guidance. The guidance is subject
to sampling error and thus a second source of noise can be introduced to the
model. This noise varies with time during the simulation and can be considered
to have two contributions; one from variations in the true expression level of
the signalling gradient and one from sampling error measuring it. We will
combine these into a single error.

@@

the term for the divergence of $\mb{J}$ can be written:
%
\begin{equation}
  \label{eq:divJ}
  \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right).
\end{equation}
%
In the barrels model, each $\rho_j(\mb{x})$ was assumed static, allowing the
definition of a time-independent $\mb{g}_i(\mb{x})$:
%
\begin{equation}
\mb{g}_i(\mb{x}) \equiv \sum_{j=1}^{M} \gamma_{i,j} \nabla\rho_j(\mb{x}).
\end{equation}
%
Here, we represent the time-dependence of the signalling molecule expression
by noting that $\rho$ is now a function of time.  The error which the growth
cones make in sampling $\rho$ and estimating $\nabla\rho_j(\mb{x})$ is
$\xi_\rho(t)$:
%
\begin{equation}
\mb{g}_i(\mb{x},t) \equiv \sum_{j=1}^{M} \gamma_{i,j} \big( \nabla\rho_j(\mb{x},t)
+ \xi_\rho(t) \big).
\end{equation}
%
One problem with this representation is that the equations represent a
population model. Within one spatial element of the discretized solution of
the differential equations, the error, $\xi_\rho(t)$ can have only one value at
time $t$; this value is applied to all the individual axon growth cones that
are represented by that element. Nonetheless, it's a useful, and fairly simple
way to represent the guidance error in the system. In practice, we ignore the
variation in $\rho_j$ and write
%
\begin{equation}
\mb{g}_i(\mb{x},t) \equiv \sum_{j=1}^{M} \gamma_{i,j} \big( \nabla\rho_j(\mb{x})
+ \xi_\rho(t) \big).
\end{equation}

@@

\subsection{Non-competition based stopping mechanism}

@@

This can be achieved with a function of the gammas and $\mb{x}$:

\begin{equation} \label{eq:da_stopping}
\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M=2} \gamma_{i,j}\nabla \rho_j(\mb{x})\right)\;f(\gamma_{i,1},\gamma_{i,2},\mb{x}) - \frac{\partial c_i}{\partial t},
\end{equation}

Where $f$ is a function which permits branching far away from a given axon's
target region and tends to zero inside the target region. It represents an
axon's internal `stopping process'. Start with a Gaussian function:

\begin{equation}\label{eq:g}
g(\mb{x}) = \mathrm{e}^{-\big(\frac{\mb{x}-\mb{\gamma}}{2\sigma}\big)^2 }
\end{equation}

where $\mb{\gamma}$ gives the location of the peak and $\sigma$ is the width
of the peak. This is squashed through an inverting sigmoid, with sharpness $s$:

\begin{equation}\label{eq:f}
f(\mb{x}) = \frac{2}{1 + \mathrm{e}^{-s g(\mb{x})}} - 1
\end{equation}

@@

\begin{equation}\label{eq:gf}
f(\mb{x}) = \frac{2}{1 + \mathrm{e}^{-s \mathrm{e}^{-\big(\frac{\mb{x}-\mb{\gamma}}{2\sigma}\big)^2 } }} - 1
\end{equation}

>>>>

On arriving at the tectum, axon growth cones must determine their direction of
movement. The model encapsulates the quantitative hypotheses that growth cone
movement is a combination of diffusion and guidance. The guidance is subject
to sampling error and thus a second source of noise can be introduced to the
model. This noise varies with time during the simulation and can be considered
to have two contributions; one from variations in the true expression level of
the signalling gradient and one from sampling error measuring it. We will
combine these into a single error. \textbf{Or maybe not:} Because the
no-competition model (see Section \ref{sec:nocomp}) has a function
$f(\gamma, \rho)$, which represents the axons' ability to sample the absolute
value of $\rho$, it is important to represent the variability of the true
value of $\rho(\mb{x})$.

@@

the term for the divergence of $\mb{J}$ can be written:
%
\begin{equation}
  \label{eq:divJ}
  \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right).
\end{equation}
%
In the barrels model, each $\rho_j(\mb{x})$ was assumed static, allowing the
definition of a time-independent $\mb{g}_i(\mb{x})$:
%
\begin{equation}
\mb{g}_i(\mb{x}) \equiv \sum_{j=1}^{M} \gamma_{i,j} \nabla\rho_j(\mb{x}).
\end{equation}
%
Here, we represent the time-dependence of the signalling molecule expression
by noting that $\rho$ is now a function of time.  The error which the growth
cones make in sampling $\rho$ and estimating $\nabla\rho_j(\mb{x})$ is
$\xi_{\rho'}(t)$:
%
\begin{equation}
\mb{g}_i(\mb{x},t) \equiv \sum_{j=1}^{M} \gamma_{i,j} \big( \nabla\rho_j(\mb{x},t)
+ \xi_{\rho'}(t) \big).
\end{equation}
%
One problem with this representation is that the equations represent a
population model. Within one spatial element of the discretized solution of
the differential equations, the error, $\xi_\rho(t)$ can have only one value at
time $t$; this value is applied to all the individual axon growth cones that
are represented by that element. Nonetheless, it's a useful, and fairly simple
way to represent the guidance error in the system. In practice, we ignore the
variation in $\rho_j$ and write
%
\begin{equation}
\mb{g}_i(\mb{x},t) \equiv \sum_{j=1}^{M} \gamma_{i,j} \big( \nabla\rho_j(\mb{x})
+ \xi_{\rho'}(t) \big).
\end{equation}

@@

\subsection{Non-competition based stopping mechanism}
\label{sec:nocomp}

@@

This can be achieved with a function of the gammas and rhos:

\begin{equation} \label{eq:da_stopping}
\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M=2} \gamma_{i,j}\nabla \rho_j(\mb{x})\right)\;f\big(\gamma_{i,1},\gamma_{i,2},\tilde{\rho}_{1}(\mb{x}),\tilde{\rho}_{2}(\mb{x})\big) - \frac{\partial c_i}{\partial t},
\end{equation}

Where $f$ is a function which permits branching far away from a given axon's
target region and tends to zero inside the target region. It represents an
axon's internal `stopping process'. Note that $\tilde{\rho}_j(\mb{x})$ is the
axon's measurement of $\rho_j(\mb{x})$ and may include measurement noise,
$\xi_\rho$: $\tilde{\rho}_j = \rho_j + \xi_\rho$. Given that the morphogen
gradients are orthogonal, $(\tilde{\rho}_1,\tilde{\rho}_2)$ defines a vector
field $\vec{\rho}(\mb{x})$. Similarly, we can combine
$(\gamma_{i,1},\gamma_{i,2})$ into the vector $\vec{\gamma}$.

This makes it easy to start to build $f$ with a two dimensional Gaussian
function:

\begin{equation}\label{eq:g}
g\big(\vec{\rho}(\mb{x})\big) = \mathrm{e}^{-\big(\frac{\vec{\rho}(\mb{x})-\vec{\gamma}}{2 w}\big)^2 }
\end{equation}

where $\vec{\gamma}$ gives the location of the peak and $w$ is the width
of the peak. This is squashed through an inverting sigmoid, with sharpness $s$:

\begin{equation}\label{eq:f}
f\big(\vec{\rho}(\mb{x})\big) = \frac{2}{1 + \mathrm{e}^{-s g(\vec{\rho}\,)}} - 1
\end{equation}

@@

\begin{equation}\label{eq:gf}
f(\vec{\rho}(\mb{x})) = \frac{2}{1
+ \mathrm{e}^{-s \mathrm{e}^{-\big(\frac{\vec{\rho}(\mb{x})-\vec{\gamma}}{2 w}\big)^2 } }} - 1
\end{equation}
